---
title: "p8105_hw1_ja3314"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

## Problem 1

*Defining my variables*

```{r}
vec_numeric = runif(10, min = 0, max = 5)
vec_logical = c(vec_numeric > 2)
vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")
vec_factor = factor(c(1,0,0,1,1,1,0,1,0,0), levels = c(0,1), labels = c("positive", "negative"))
```

*Creating data frame using variables*

```{r}
uniform_dist_df = tibble(
  vec_numeric,
  vec_logical,
  vec_char,
  vec_factor,
)
```

*Calculating the mean of each variable in uniform_dist_df:*

```{r}
summarise_at(uniform_dist_df, c("vec_numeric", "vec_logical", "vec_char", "vec_factor"), mean)
```

The mean calculation only works for the numeric and logical vectors. The character and factor vectors do not contain numeric values and hence a mean cannot be calculated. The logical vector values are coerced to numeric and therefore a value has been calculated.

*Converting logical, character, and factor vectors into numeric vectors:*

```{r, eval = FALSE}
as.numeric(vec_logical)
as.numeric(vec_char)
as.numeric(vec_factor)
```

The logical and factor vector are able to be converted to numeric values but the character vector is not.

*Converting character variable from character to factor to numeric and factor variable from factor to character to numeric*

```{r}
char_factor_numeric = as.factor(vec_char)
as.numeric(char_factor_numeric)

factor_char_numeric = as.character(vec_factor)
as.numeric(factor_char_numeric)
```

Converting a character vector to a factor vector and then using the `as.numeric` function produces a numeric factor however converting a factor vector to a character vector does not.


