---
title: "p8105_hw1_ja3314"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Loading the tidyverse package
library(tidyverse)

```

## Problem 1

**Creating data frame**
```{r}
uniform_dist_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5), 
  vec_logical = c(vec_numeric > 2), 
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"), 
  vec_factor = factor(c(1,0,0,1,1,1,0,1,0,0), levels = c(0,1), labels = c("positive", "negative"))
)
```

**Calculating the mean of each variable in uniform_dist_df:**
```{r, warning=FALSE}
summarise_at(uniform_dist_df, c("vec_numeric", "vec_logical", "vec_char", "vec_factor"), mean)
```
The mean calculation only works for the numeric and logical vectors. The character and factor vectors do not contain numeric values and hence a mean cannot be calculated. The logical vector values are coerced to numeric and therefore a value has been calculated.


**Converting logical, character, and factor vectors into numeric vectors:**
```{r, eval = FALSE}
as.numeric(uniform_dist_df$vec_logical)
as.numeric(uniform_dist_df$vec_char)
as.numeric(uniform_dist_df$vec_factor)
```

The logical and factor vector are able to be converted to numeric values but the character vector is not.

**Converting character variable from character to factor to numeric and factor variable from factor to character to numeric**

```{r, warning=FALSE}
char_factor_numeric = as.factor(uniform_dist_df$vec_char)
as.numeric(char_factor_numeric)

factor_char_numeric = as.character(uniform_dist_df$vec_factor)
as.numeric(factor_char_numeric)
```

Converting a character vector to a factor vector and then using the `as.numeric` function produces a numeric factor however converting a factor vector to a character vector does not.


## Problem 2

**Creating data frame:**

```{r}

norm_dist_df = tibble(
  x = rnorm(1000), 
  y = rnorm(1000), 
  log_vec = x + y > 0, 
  num_vec = as.numeric(log_vec), 
  fac_vec = as.factor(log_vec), 
)
```


```{r, include=FALSE}
## Calculating descriptive measures for norm_dist_df
dim_norm_dist_df = dim(norm_dist_df)
mean_x = mean(norm_dist_df[["x"]])
median_x = median(norm_dist_df[["x"]])
num_true = table(norm_dist_df[["log_vec"]])["TRUE"]
prop_true = (num_true/1000)
```

This vector's dimensions are [`r dim_norm_dist_df`]. x has a mean of `r mean_x` and a median of `r median_x`. The proportion of `TRUE` cases in the logical vector is `r prop_true`. 

**Creating scatterplots:**

*Color-coded based on logical variable:*
```{r, message=FALSE}
ggplot(norm_dist_df, aes(x = y, y = x, color = log_vec)) + geom_point()

ggsave("norm_dist_plot.pdf")
```

*Color-coded based on numeric variable:*
```{r}
ggplot(norm_dist_df, aes(x = y, y = x, color = num_vec)) + geom_point()
```

*Color-coded based on factor variable:*
```{r}
ggplot(norm_dist_df, aes(x = y, y = x, color = fac_vec)) + geom_point()
```

Since the values for my logical and factor variables are the same, their plots are also color-coded the same way. Even though the numeric variable only has 2 possible values, the color-coding for the scatterplot is a gradient since there is a potential for more than 2 variables in a numeric vector.